---
title: "Homework 3"
author: "Xiangyi Liu (xl3048)"
date: "10/6/2020"
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1
```{r data}
data("instacart")
```
This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

Make plot.
```{r scatterplot}
instacart %>%
  count(aisle) %>%
  filter(n > 1000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle,n) #reorder factor 'aisle' based on 'n'
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1) )
```

Make a table.
```{r table}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care","packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

Apple vs. ice cream.
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apple", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )

```

## Problem 2
```{r data import}
accel_df <-
  read.csv("accel_data.csv") %>%
  rename(day_of_week = day) 
  
day_info <-
  accel_df %>% select(week:day_of_week)

accel_df <-
  accel_df%>%
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute_of_day",
    names_prefix = "activity.",
    values_to = "activity_count"
  ) %>%
  mutate(
    minute_of_day = as.numeric(minute_of_day),
    day_id = as.factor(day_id),
    week = as.factor(week),
    day_of_week = as.factor(day_of_week),
    weekday_weekend = case_when(
      day_of_week %in% c("Saturday","Sunday") ~ "weekend",
      day_of_week %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday"
    ) ,
    weekday_weekend = as.factor(weekday_weekend)
  ) %>%
  relocate(week,day_id,weekday_weekend)
```

The resulting dataset has `r nrow(accel_df)` observations and `r ncol(accel_df)` variables, which are {`r colnames(accel_df)`}. All variables are factors except minute_of_day and activity_count, which is numeric.

```{r, message=FALSE}
accel_total <-
  accel_df %>%
  group_by(day_id) %>%
  summarise(total_activity = sum(activity_count)) 

day_info <-
  day_info %>%
  mutate(
    weekday_weekend = case_when(
      day_of_week %in% c("Saturday","Sunday") ~ "weekend",
      day_of_week %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday"
    ) ,
    weekday_weekend = as.factor(weekday_weekend),
    day_id = as.factor(day_id)
  )

day_info <-
 day_info %>% left_join(accel_total, by = "day_id") 
  
day_info %>% knitr::kable()
```
From the table alone, we can observe that the person gradually has less activity when the recording approaching to the end. There are two extreme low values on day 24 and day 31. On average, the person has less activity on Wednesdays.

```{r plot_p2}
ggplot(accel_df, aes(x=minute_of_day, y=activity_count,color=day_id)) +
  geom_point(alpha = 0.5) +
  labs(title = "Activity Over The Day",
       x = "Minute of the day",
       y = "Activity Count") +
  scale_x_continuous(breaks = round(seq(0,1440,by=120)))
```

```{r}
ggplot(accel_df, aes(x=minute_of_day, y=activity_count,color=day_id)) +
  geom_smooth(se=FALSE,method = "lm")+
  labs(title = "Activity Over The Day",
       x = "Minute of the day",
       y = "Activity Count") +
  scale_x_continuous(breaks = round(seq(0,1440,by=120)))
```

For clarity, the scatterplot and their trendlines are plotted seperately. 
The person tend to have more activity at the beginning of the study, and graduallt decreasing as the study approached to its end.


## Problem 3 
```{r}
data("ny_noaa")
ny_noaa <-
  ny_noaa %>%
  na.omit() %>%
  separate(date, sep = "-", into = c("year","month","day")) %>%
  mutate(
    month = as.numeric(month),
    month = month.name[month],
    tmin = as.numeric(tmin)/10,
    tmax = as.numeric(tmax)/10
  )

snow_rank<-
  plyr::count(ny_noaa,"snow") %>%
  arrange(desc(freq))

head(snow_rank)%>%knitr::kable()

```
The frequency of each unique observed value for 'snow' is counted and sorted in a decreasing order. The five most commonly observed values are shown in the table above.

```{r}
jan_tmax<-
  subset(ny_noaa,month=="January") %>%
  group_by(id,year) %>%
  summarize(
    mean_tmax=mean(tmax)
  )%>%
  mutate(
    month = "January"
  )%>%
  relocate(month,everything())

jul_tmax<-
  subset(ny_noaa,month=="July") %>%
  group_by(id,year) %>%
  summarize(
    mean_tmax=mean(tmax)
  )%>%
  mutate(
    month = "July"
  )%>%
  relocate(month,everything())

comp_tmax<-
  rbind(jan_tmax,jul_tmax) %>%
  na.omit()%>%
  mutate(
    year=as.factor(year)
  )
```

```{r plot_janvsjul}
ggplot(comp_tmax, aes(x=year,y=mean_tmax,color=month))+
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ month)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1) )+
  labs(title = "January vs. July")
```
The average max temperature for each year in January ranges from `r quantile(jan_tmax$mean_tmax,0.25)` to 
`r quantile(jan_tmax$mean_tmax,0.75)`; The average max temperature for each year in July ranges from `r quantile(jul_tmax$mean_tmax,0.25)` to 
`r quantile(jul_tmax$mean_tmax,0.75)`. Moreover, the average max temperature for January varies more across stations in the same year.
From the plot we could observe that the max temperature in January has more outliers compared to July. There are a couple of noticeable outliers, such as the January in 1982 and 1996, the July in 1988.

```{r hex}
library(hexbin)
temp_hex<-ggplot(ny_noaa, aes(x = tmax, y = tmin)) + 
  geom_hex()+
  labs(title = "tmin & tmax")
```

```{r}
filtered<-
  ny_noaa %>%
  filter((snow>0))%>%
  filter(snow<100)%>%
  mutate(year = forcats::fct_reorder(year, snow))

snow_dist<-
  ggplot(filtered, aes(x = year,y=snow)) + 
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1) )+
  labs(title="Snowfall Distribution")
```

```{r combine}
library(patchwork)
temp_hex+snow_dist
```




